Optimal (Multiway) Spatial Joins

### 2024-05-24 msj.tex
yufei

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\out{\mathrm{OUT}}

@@

\extraspacing {\bf Generating $\bm{\J^*}$.} Take an arbitrary $(k-2)$-tuple $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2}) \in \J(R_1', R'_2, ..., R'_{k-2})$. The reader should recall from Section~\ref{sec:bricks} that
\myitems{
    \item $B_\bm{t}$ is $\bigcap_{i=1}^{k-2} \bm{t}[i]$, i.e., the intersection of $r'_1, r'_2, ..., r'_{k-2}$;
    \item $\gleft(\bm{t})$ is the $r'_i$ ($1 \le i \le k-2$) with $\xleft(r'_i) = \xleft(B_\bm{t})$;
    \item $\gbot(\bm{t})$ is the $r'_i$ ($1 \le i \le k-2$) with $\ybot(r'_i) = \ybot(B_\bm{t})$.
}
We now introduce:
% \myitems{
%     \item $\cross(r_a, B; H)$, the set of segments in $H$ that cross $r_a$ and $B$;
%     \item $\cross(r_b, B; V)$, the set of segments in $V$ that cross $r_b$ and $B$.
% }
\myeqn{
    \cross(\bm{t}, H) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}} \nn \\
    \cross(\bm{t}, V) = \set{h \in H \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \nn
}
These sets have important properties as stated in the next lemma, whose proof can be found in Appendix~\ref{app:hv}:

@@

        \myenums{
            \item $\cross(\bm{t}, H) \ne \emptyset$ and $\cross(\bm{t}, V) \ne \emptyset$.

            \item Denote by $r_i$ ($i \in [k-2]$) the full rectangle of $\bm{t}[i]$. 
            For any $h \in \cross(\bm{t}, H)$ and any $v \in \cross(\bm{t}, V)$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.

            \item If the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $h \in \cross(\bm{t}, H)$ and $v \in \cross(\bm{t}, V)$.
        }

@@

        \item $\sum_{\bm{t}} |\cross(\bm{t}, H)| \le \out$ and $\sum_{\bm{t}} |\cross(\bm{t}, V)| \le \out$.
    }
\end{lemma}

@@

\mytab{
    \> {\bf algorithm generate-$\J^*$} \\
    \> 1.\> $\J^* = \emptyset$ \\
    \> 2.\> {\bf for} each $(k-2)$-tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ {\bf do} \\
    \> 3.\>\> $r_i \leftarrow$ the full rectangle of $\bm{t}[i]$, for each $i \in [k-2]$ \\
    \> 4.\>\> {\bf for} each $(h, v) \in \cross(\bm{t}, H) \times \cross(\bm{t}, V)$ {\bf do} \\
    \> 5.\>\>\> add $(r_1, ..., r_{k-2}, h, v)$ to $\J^*$
}

Lemma~\ref{lmm:hv:type1:properties} assures us that the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}. Furthermore, if we are given $\cross(\bm{t}, H)$ and $\cross(\bm{t}, V)$ for each $\bm{t}$, Lemma~\ref{lmm:hv:type1:properties} further asserts that the above algorithm runs in $O(1 + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time.

@@

The rest of the section will focus on explaining how to prepare the sets $\cross(\bm{t}, H)$ of all $\bm{t} \in \J(R_1', ..., R_{k-2}')$ in $O(n \log n + k \cdot \out)$ time. An analogous algorithm can be used to compute the sets $\cross(\bm{t}, V)$ of all $\bm{t}$ in the same time complexity.

@@

Now construct from $R_i'$ a new set of rectangles: 
\myeqn{
    R^*_i = \set{r^* \mid r' \in R_i' \text{ and its top-sliced rectangle $r^*$ exists}}. \nn
}

@@

>>>>

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\dcross{\mathrm{d\text{-}cross}}
\def\out{\mathrm{OUT}}

@@

\extraspacing {\bf Generating $\bm{\J^*}$.} Take an arbitrary $(k-2)$-tuple $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2}) \in \J(R_1', R'_2, ..., R'_{k-2})$. The reader should recall from Section~\ref{sec:bricks} that
\myitems{
    \item $B_\bm{t}$ is $\bigcap_{i=1}^{k-2} \bm{t}[i]$, i.e., the intersection of $r'_1, r'_2, ..., r'_{k-2}$;
    \item $\gleft(\bm{t})$ is the $r'_i$ ($1 \le i \le k-2$) with $\xleft(r'_i) = \xleft(B_\bm{t})$;
    \item $\gbot(\bm{t})$ is the $r'_i$ ($1 \le i \le k-2$) with $\ybot(r'_i) = \ybot(B_\bm{t})$.
}
We now introduce:
% \myitems{
%     \item $\cross(r_a, B; H)$, the set of segments in $H$ that cross $r_a$ and $B$;
%     \item $\cross(r_b, B; V)$, the set of segments in $V$ that cross $r_b$ and $B$.
% }
\myeqn{
    \dcross(\bm{t}, H) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}} \nn \\
    \dcross(\bm{t}, V) = \set{h \in H \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \nn
}
The prefix ``d-'' stands for double. These sets have important properties as stated in the next lemma, whose proof can be found in Appendix~\ref{app:hv}:

@@

        \myenums{
            \item $\dcross(\bm{t}, H) \ne \emptyset$ and $\dcross(\bm{t}, V) \ne \emptyset$.

            \item Denote by $r_i$ ($i \in [k-2]$) the full rectangle of $\bm{t}[i]$. 
            For any $h \in \dcross(\bm{t}, H)$ and any $v \in \dcross(\bm{t}, V)$, the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J(R_1, ..., R_{k-2}, H, V)$.

            \item If the $k$-tuple $(r_1, r_2, ..., r_{k-2}, h, v)$ belongs to $\J_1$, then $h \in \dcross(\bm{t}, H)$ and $v \in \dcross(\bm{t}, V)$.
        }

@@

        \item $\sum_{\bm{t}} |\dcross(\bm{t}, H)| \le \out$ and $\sum_{\bm{t}} |\dcross(\bm{t}, V)| \le \out$.
    }
\end{lemma}

@@

\mytab{
    \> {\bf algorithm generate-$\J^*$} \\
    \> 1.\> $\J^* = \emptyset$ \\
    \> 2.\> {\bf for} each $(k-2)$-tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ {\bf do} \\
    \> 3.\>\> $r_i \leftarrow$ the full rectangle of $\bm{t}[i]$, for each $i \in [k-2]$ \\
    \> 4.\>\> {\bf for} each $(h, v) \in \dcross(\bm{t}, H) \times \dcross(\bm{t}, V)$ {\bf do} \\
    \> 5.\>\>\> add $(r_1, ..., r_{k-2}, h, v)$ to $\J^*$
}

Lemma~\ref{lmm:hv:type1:properties} assures us that the set $\J^*$ thus computed indeed satisfies \eqref{eqn:hv:type1:J*}. Furthermore, if we are given $\dcross(\bm{t}, H)$ and $\dcross(\bm{t}, V)$ for each $\bm{t}$, Lemma~\ref{lmm:hv:type1:properties} further asserts that the above algorithm runs in $O(1 + k \cdot |\J^*|) = O(1 + k \cdot \out)$ time.

@@

The rest of the section will focus on explaining how to prepare the sets $\dcross(\bm{t}, H)$ of all $\bm{t} \in \J(R_1', ..., R_{k-2}')$ in $O(n \log n + k \cdot \out)$ time. An analogous algorithm can be used to compute the sets $\dcross(\bm{t}, V)$ of all $\bm{t}$ in the same time complexity.

@@

Now construct from $R_i'$ a new set of rectangles: 
\myeqn{
    R^*_i = \set{r^* \mid r' \in R_i' \text{ and its top-sliced rectangle $r^*$ exists}}. \label{eqn:hv:type1:R*}
}
In Appendix~\ref{app:hv}, we show how to compute produce $R^*_1, ..., R^*_{k-2}$ together in $O(n + k \cdot \out)$ total time.

@@

\extraspacing {\bf Computing the Top-Sliced Rectangles.}
